### 2016-04-07 paper/paper.tex
Revision after review.

\documentclass{llncs}
\usepackage{amsmath,amsfonts,amssymb}
\usepackage{graphicx}
\usepackage{makeidx}  % allows for indexgeneration
\begin{document}

\title{Simulation of Tsunami impact upon Coastline}

\titlerunning{Tsunami simulation}

\author{Aristotelis Spathis-Papadiotis\inst{1} \and Konstantinos Moustakas\inst{1}}
\authorrunning{Aristotelis Spathis-Papadiotis et al.} % abbreviated author list (for running head)

%%%% list of authors for the TOC (use if author list has to be modified)
\tocauthor{Konstantinos Moustakas, Aristotelis Spathis-Papadiotis}

\institute{Electrical and Computer Engineering Department\\
  University of Patras, Patra, Greece\\
  \email{moustakas@upatras.gr}}

\maketitle

\begin{abstract}
  This paper presents a simulation of a tsunami impact upon an urban coastline. Emphasis
  was given to the conservation of momentum, as its distribution in space and time is the
  main factor of the wave's effects on the coastline. Due to this, a hybrid simulation
  method was adopted, based on the Smoothed Particle Hydrodynamics (SPH) method, enriched
  with geometric constraints and rigid body interactions. The implementation is the result
  of cooperation between the Bullet physics engine and our custom SPH engine, which
  successively process the dynamic state of the fluid at every timestep. Furthermore, in
  order to achieve better performance a custom data structure (LP grid) was developed for
  the optimization of locality in data storage and minimization of access time. Simulation
  data is exported to VTK files, allowing interactive processing and
  visualization. Experimental results demonstrate the benefits of impulse recording at
  potential hazard estimation and evaluation of defense strategies. \keywords{Fluid
    simulation, tsunami, SPH, tsunami-coastline interaction, force visualization}
\end{abstract}

\section{Introduction}

\paragraph{} Simulations of natural phenomena are a precious tool for analysis and
understanding of the processes behind them as well as the implications of
those. Especially fluid dynamics is one of the fields most benefitted by the explosive
growth of high performance parallel computing architectures of the last years. Tsunamis
are one of the most devastating natural disasters, with much attention drawn to them
lately, especially after the 2004 Indian Ocean tsunami and the 2011 T\={o}hoku earthquake
and tsunami, two of the largest incidents in modern history. Multiscale modelling of
tsunami generation, propagation and impact is a trending research area as respective
simulations give valuable insights into the underlying mechanisms and relations between
the various stages of an unfolding tsunami incident, while also facilitating the
assessment of potential hazard it poses upon impact on a coastline.

\paragraph{} A tsunami is a series of waves in a water body caused by an impulsive
disturbance that vertically displaces a large volume of water. Tsunamis are generated by
earthquakes, volcanic eruptions, landslides and other such events which have the potential
to transmit a huge amount of mechanical energy to an overlying or adjacent water
volume. On a macroscopic level, tsunami propagation across the ocean as well as coastline
inundation and runup are usually simulated through methods based on various versions of
the shallow water equations, which are derived from the Navier-Stokes equations if the
horizontal length scale is much greater than the vertical one. Conversely, the
aforementioned methods are not appropriate in smaller scale simulations of the impact upon
the coastline, since in order to obtain a reliable estimation of the forces exerted on the
terrain, complex dynamics of fluid-structure interactions have to be accounted for. One of
the most used methods for simulating complex flows with multiple boundary interactions is
Smoothed Particle Hydrodynamics, initially proposed in the 1970s for the treatment of
compressible flows in astrophysical problems. The method has enjoyed extensive adoption,
having been applied to numerous fields ranging from aerodynamics and geology to
engineering and computer graphics with exceptional results, lending itself well to
adaptations and extensions. The main contribution of our approach is the emphasis given to
the explicit conservation of momentum and the detailed data collection and visualization
relating to its distribution upon the coastline during the tsunami impact. Towards this
end, we adopted SPH enriched with geometric constraints as our method of choice due to the
unparalleled advantages it offers, directly linked to its properties as a lagrangian
method.

\section{Related Work}

\paragraph{} Computational fluid dynamics is a very active area of research, as it relates
to a wide spectrum of applications in computer graphics, engineering and science. The SPH
method was developed independently by Gingold and Monaghan \cite{gingold1977375} and Lucy
\cite{lucy19771013} in 1977. Since then, it has been employed in a wide variety of
problems and applications, proving itself to be a flexible and attractive method for the
simulation of complex, multicomponent/multiphase flows. Its flexibility is shown by the
numerous adaptations and customizations it has undergone to suit many diverse problem
domains.

\paragraph{} Tsunami simulations have been shown to be a valuable tool in the literature,
aiding in estimation of key parameters of a tsunami incident, such as its propagation
course and coastal inundation. Simulations of the devastating 2004 Indian Ocean tsunami
have been carried out by Wang and Liu \cite{wang2007} and Ioualalen et
al. \cite{ioualalen2007}, using nonlinear shallow water equations and nonlinear Boussinesq
models respectively. In both cases, good agreement between calculations and field
measurements of inundation and runup indicate the robustness of these approches. Samaras
et al. \cite{samaras2015} used a 2D horizontal model based on the higher oder Boussinesq
equations to simulate tsunami-induced coastal inundation in two relevant areas of interest
in the Mediterranean. Using two different models, Kakinuma \cite{kakinuma2008} reproduced
a past incident with sufficient accuracy. The governing equations for both were the
continuity and Reynolds equations for incompressible fluids in porous media, while water
surface displacement was determined by the vertically integrated continuity equation for
the first and the 3D Volume of Fluid method for the second model.

\paragraph{} Relating to the SPH method, Desbrun and Gascuel \cite{desbrun1996smoothed}
used it to animate highly deformable bodies of various stiffness and viscosity, while
proposing important extensions like the ``spiky'' pressure smoothing kernel and discussing
implementation issues such as the fluid surface reconstruction from density
isosurfaces. Later, M\"{u}ller et al. \cite{muller2003particle} used this work as a basis
for fluid simulation in interactive applications. Becker and Teschner
\cite{becker2007weakly} developed Weakly Compressible SPH, where the ideal gas equation of
state is replaced with the much more strict Tait equation to reduce compressibility to a
user-defined upper bound, thus avoiding an inefficient explicit Poisson equation
solver. In recent years, Solenthaler and Pajarola \cite{solenthaler2009predictive}
improved WCSPH by proposing Predictive-Corrective Incompressible SPH, where a
prediction-correction scheme is used to compute particle pressures through iterative
density constraint satisfaction and propagation through the fluid, until the final
tolerance conditions are met. Implicit Incompressible SPH was then proposed by Ihmsen et
al. \cite{ihmsen2014implicit}, in which the density is predicted using a discretized form
of the pressure Poisson equation, as obtained from the combination of a direct
discretization of the continuity equation and symmetric pressure forces.

\paragraph{} SPH has been employed in many simulations involving wave-coastline
interactions. Debroux et al. \cite{debroux2001three} used SPH for the simulation of
tsunami impact on real coastline topography, where important features like the energy and
speed of the wave were measured over the course of time. G\'{o}mez-Gesteira and Dalrymple
\cite{gomez2004} simulated the impact of a wave generated by a dam break against a tall
structure with a 3D SPH method, showing good agreement with lab measurements of velocities
and forces. Dalrymple and Rogers \cite{dalrymple2006} gave a comprehensive overview of
techniques regarding usage of SPH for simulation of water waves and their interaction with
structures, while also highlighting its suitability for close-up examinations of small
regions. Many studies have considerably improved upon this line of research, yielding
exceptional results. For example, St-Germain et al. \cite{st-germain2014} used a 3D WCSPH
model to examine and analyze spatially and temporally forces exerted by tsunami-like
hydraulic bore upon freestanding structures.

\paragraph{} Due to its generality and adaptability, SPH has been the method of choice for
many applications and relative extensions. Macklin and M\"{u}ller
\cite{macklin2013position} enriched SPH with positional geometric constraints, thus
enforcing incompressibility while maintaining stability and allowing for large
timesteps. Rigid-fluid interaction is one of the strong aspects of the method, as is shown
by Akinci et al. \cite{akinci2012versatile}, where rigid surfaces are sampled by boundary
particles, which adaptively contribute to fluid properties to address boundary region
deficiencies and inhomogeneities, in order to incorporate rigid bodies to a unified
hydrodynamic framework. Finally, a thorough overview of various implementation algorithms
and data structures together with their advantages and drawbacks is provided by
Dom\'{i}nguez et al. \cite{dominguez2011}.

\section{Proposed framework}

\paragraph{} Our proposed simulation framework is based on SPH, a lagrangian fluid
simulation method whose core notion is the discretization of the fluid into particles,
which serve as interpolation points for the estimation of fluid properties in
space. Advantages of this method include the exact treatment of advection, the natural way
of dealing with special interface interactions, the inherent conservation of significant
quantities (mass, momentum, energy) and the self-adaptivity of computational load to the
fluid location and state in the flow domain. To efficiently ensure incompressibility in
degenerate cases and undersampled boundary regions while maintaining relatively large
timesteps, SPH was enhanced with explicit solving of geometric constraints between
particles. These constraints are solved by the Bullet physics engine, within which fluid
particles are represented as rigid bodies, while also being subject to SPH forces as
computed at each timestep of the simulation.

\subsection{SPH method}

\paragraph{} Starting from the identity:
\begin{equation*}
  f(\vec{r}) = \int_Vf(\vec{x}) \delta(\vec{r} - \vec{x}) d\vec{x},
\end{equation*}
where $\delta(\vec{r})$ the Dirac delta function and $\vec{x} \in V$, one can obtain a
more general interpolation rule by substituting $\delta(\vec{r})$ with a smoothing kernel
$W(\vec{r}, h)$:
\begin{equation*}
  \label{eq:c-approx}
  f(\vec{r}) \approx \int_V f(\vec{x}) W(\vec{r}-\vec{x}, h) d\vec{x}
\end{equation*}
whose limit when $h\to0$ approaches the delta fucntion and is normalized to unity:
\begin{equation*}
  \label{eq:kernel-properties}
  \lim_{h\to0}W(\vec{r}, h) = \delta(\vec{r})\ \ \
  \text{and}\ \ \
  \int_VW(\vec{r}, h) d\vec{x} = 1
\end{equation*}
The smoothing radius $h$ serves as a cutoff radius in the smoothing process, as particles
beyond that distance have no contribution to the sum, i.e. $W(r,h) = 0$ when $r>h$. For
the discrete case, where $f$ is discretized to particles with density $\rho$ and mass $m$,
the weighting ratio $m/\rho$ can be used to construct a weighted sum interpolant for any
field $A$:
\begin{equation}
  \label{eq:d-approx}
  A(\vec{r}) = \sum_i \frac{m_i}{\rho_i} A(\vec{r}_i) W(\vec{r}-\vec{r}_i, h)
\end{equation}
which lies at the heart of SPH formulation. According to this, the gradient can be
computed by the following approximation:
\begin{equation}
  \label{eq:d-grad}
  \nabla A(\vec{r}) = \sum_i \frac{m_i}{\rho_i} A(\vec{r}_i) \nabla W(\vec{r} - \vec{r}_i, h)
\end{equation}
The obvious advantage of this is the exclusive dependence on the smoothing kernel
gradient, which can be precomputed for sensible kernel choices. However, this formula can
lead to unsymmetric pair forces, compromising the conservation of linear and angular
momentum of the system. To symmetrize these forces depending on gradients (like those
originating from pressure differences), we can use the product rule:
\begin{equation*}
  \label{eq:grad-identity}
  \nabla \left( \frac{P}{\rho} \right) =
  \frac{\nabla P}{\rho}-
  \frac{P}{\rho^2} \nabla \rho
  \hspace{10pt} \Leftrightarrow \hspace{10pt}
  \nabla P = \rho \left[ \frac{P}{\rho^2} \nabla \rho + \nabla \left( \frac{P}{\rho}
    \right) \right]
\end{equation*}
to obtain an alternative approximation of gradient
\begin{align}
  \label{eq:grad-est}
  \nabla P & = \rho \left[ \frac{P}{\rho^2} \sum_i \frac{m_i}{\rho_i} \rho_i \nabla
             W(\vec{r}-\vec{r}_i, h)
             +
             \sum_i \frac{m_i}{\rho_i} \frac{P_i}{\rho_i} \nabla W(\vec{r}-\vec{r}_i, h)
             \right] \nonumber \\
           & = \rho \sum_i m_i \left(\frac{P}{\rho^2} + \frac{P_i}{\rho_i^2} \right)
             \nabla W(\vec{r}-\vec{r}_i, h)
\end{align}
which is antisymmetric for all interacting particle pairs. Viscosity forces on the other
hand, are proportional to the laplacian of the velocity field:
\begin{equation}
  \label{eq:lapl-est}
  \nabla^2\vec{v} = \sum_i \frac{m_i}{\rho_i} (\vec{v}_i - \vec{v}) \nabla^2
  W(\vec{r}-\vec{r}_i, h)
\end{equation}
and are always antisymmetric, since they depend on velocity difference
$\vec{v}_i - \vec{v}$ between particles.  In each timestep of the simulation, the density
of all particles is first computed according to equation (\ref{eq:d-grad}), as it depends
only on the relative position of those. The pressure at each particle location is then
obtained from its respective density through an equation of state. After this step, the
pressure and viscosity forces are computed from the particle data and integrated back into
the position and velocity of the particles.

\subsection{Implementation}

\subsubsection{Simulation initialization} Each simulation under our implementation
consists of two elements, the coastline terrain (static) and the tsunami wave
(dynamic). For the simulation setup, terrain is imported from a suitable 3D geometry
definition file format, the initial conditions of the impacting wave (position and
velocity) are configured and the desired discretization resolution for the fluid (i.e. the
number of particles it is discretized into) is set. From these conditions, the terrain and
fluid are initialized and key parameters are computed as subsequently described. Terrain
geometry is uniformly scaled by a user-supplied factor and docked to the origin of
coordinates and the particle effective radius $r$ is determined such that the specified
discretization resolution is achieved. Fluid particles are initially placed on a Hexagonal
Close-Packed lattice covering the fluid initial position, in order to accomplish the
densest possible packing and symmetry:
\begin{equation}
  \label{eq:hcp}
  [x, y, z] = \left[ 2i+[(j+k) \bmod 2], \sqrt{3}[j + \frac{1}{3} (k \bmod 2)],
    \frac{2\sqrt{6}}{3} k \right] r
\end{equation}
In this configuration, the smoothing radius $h$ is computed such that each particle has
approximately 50 neighbours (following the empirical rule established in the
literature). The timestep is calculated according to the Courant-Friedrichs-Lewy
criterion:
\begin{equation}
  \label{eq:cfl}
  \delta t_{\text{CFL}} = C \frac {\delta x}{v_{\text{max}}},
\end{equation}
for values of Courant number $C\approx0.5$ with characteristic length $\delta x$ equal to
the particle effective radius and maximum velocity $v_{\text{max}}$ determined by the
maximum potential energy in the initial configuration.

\subsubsection{SPH parameters} Following standard practice, we used three different
smoothing kernels for density, pressure gradient and velocity laplacian computation,
respectively:
\begin{equation}
  \label{eq:poly6}
  W_{\text{poly6}}(r, h) = \frac{315}{64 \pi h^9}
  \begin{cases}
    (h^2 - r^2)^3 & 0 \leq r \leq h\\
    0 & \text{otherwise}
  \end{cases}
\end{equation}
\begin{equation}
  \label{eq:spiky}
  W_{\text{spiky}}(r, h) = \frac{15}{\pi h^6}
  \begin{cases}
    (h - r)^3 & 0 \leq r \leq h\\
    0 & \text{otherwise}
  \end{cases}
\end{equation}
\begin{equation*}
  \label{eq:spiky-gradient}
  \nabla W_{\text{spiky}}(r, h) = \frac{-45}{\pi h^6} (h-r)^2
\end{equation*}
\begin{equation}
  \label{eq:viscosity}
  W_{\text{viscosity}}(r, h) = \frac{15}{2 \pi h^3}
  \begin{cases}
    - \frac{r^3}{2h^3} + \frac{r^2}{h^2} + \frac{h}{2r} - 1 & 0 \leq r \leq h\\
    0 & \text{otherwise},
  \end{cases}\\
\end{equation}
\begin{equation*}
  \label{eq:viscosity-laplacian}
  \nabla^2 W_{\text{viscosity}}(r, h) = \frac{45}{\pi h^6} (h-r)
\end{equation*}
In each simulation step iteration, $W_{\text{poly6}}$ is used as smoothing kernel for the
fluid density estimation at each particle, based on its neighbour particle
locations. $W_{\text{spiky}}$ and $W_{\text{viscosity}}$ are substituted into equations
(\ref{eq:grad-est}) and (\ref{eq:lapl-est}), in order to compute forces arising from
pressure differences and viscosity, respectively. Fluid pressure is obtained from the
estimated fluid density through the ideal gas equation of state:
\begin{equation}
  \label{eq:ideal-state}
  P = k(\rho - \rho_0),
\end{equation}
according to which the pressure is proportional to the difference of the current from the
rest density. The major problem with this equation of state are the compressibility issues
that have been shown to exist in simulations using it. A frequently proposed solution is
to replace the above with the Tait equation of state:
\begin{equation}
  \label{eq:tait-state}
  P = B \left( \left( \frac{\rho}{\rho_0} \right)^\gamma - 1 \right),
\end{equation}
where usually $\gamma=7$ and $B$ is a proportionality constant controlling the tolerance
for density fluctuations. This equation is much more punishing on density fluctuations
away from the rest density, therefore requiring significantly smaller timesteps to ensure
stability.

\subsubsection{Enhancements to SPH} In our framework we opted for a different solution,
inspired by Position Based Fluids \cite{macklin2013position}, in which particles are
represented by spherical rigid bodies. The most important advantage of this technique is
the elegant handling of boundary, undersampled and degenerate fluid regions that tend to
arise very frequently in simulations of free flows. The adoption of this approach allows
to treat boundary collisions in a simple manner, while naturally enforcing
incompressibility near boundaries and free surfaces. In these regions, estimators fail to
describe the actual flow regime due to neighbour particle shortage. This undersampling
creates the need for correction procedures, usually involving ghost/boundary particles, to
avoid pressure instabilities, particle clustering and other artifacts. On the contrary, no
such method is necessary under our representation, where degenerative cases are handled
through geometric constraints, thus selectively preventing unreliable estimations from
affecting the flow.

\begin{figure}
  \includegraphics[width=\textwidth]{../report/figures/lp-grid.pdf}
  \caption{Organization of the custom data structure designed for the efficient storing
    and interaction scanning between simulation particles. Top-down, the first vector
    implements the 3D to linear locality preserving mapping of cell anchors onto the
    second, which in turn points to each cell's particles on the third.}
  \label{fig:lp-grid}
\end{figure}

\subsubsection{Data organization and export} Rigid body dynamics are handled by the Bullet
physics engine. A custom cell list data structure (LP grid) was implemented in order to
optimize particle dynamic storage and access time. The simulation domain is divided by a
regular 3D grid of spacing equal to the smoothing radius into cubic cells containing the
fluid particles. Particles are generated as Bullet objects and are stored in the LP grid,
which consists of three storage vectors. As shown in Figure \ref{fig:lp-grid}, particles
are accessed by following pointers through those vectors in order. The \texttt{map} vector
contains pointers to \texttt{anchors} and in turn \texttt{anchors} pointers to
\texttt{particles}, which contains the simulation particles. Each cell's particles are
continuously stored in the \texttt{particles} vector. Each element of \texttt{anchors}
represents a cell and contains a pointer to the start of the region that contains that
cell's particles in \texttt{particles}. Since SPH access patterns are heavily reliant on
neighbour searching, increasing the locality in the linear order in which cells are stored
is clearly beneficial to performance, due to better utilization of cache memory. For that
reason, cells are represented spatially sorted along a space-filling curve in
\texttt{anchors} (and consequently, \texttt{particles}), in order to maximize locality
preservation from the 3D simulation domain to the linear storage. To efficiently address
the cells, one extra layer encoding the locality preserving mapping has to be added, in
the form of the first vector \texttt{map}. To access a cell, the three indices $i_x$,
$i_y$, $i_z$ specifying its position in the grid are converted to a linear index $i$ (a
simple way is through the formula $i = i_z * N_y * N_x + i_y * N_x + i_x$, where $N_x$
$N_y$ $N_z$ is the number of cells along the $x$, $y$ and $z$ axis in the grid
respectively). The \texttt{map} vector serves as a lookup table, so that its $i$-th
element contains the pointer to that cell's representation on \texttt{anchors}. LP grid
advantages include cohesive storage, quick neighbour search and interaction scanning,
exploitation of the SPH algorithm access patterns and fast, in-place update.

\paragraph{} Simulation is following the Bullet framework, with the SPH code being
embedded as an internal timestep tick callback function. We took advantage of the Bullet
infrastructure to extract detailed information about the collisions between fluid and
terrain regarding the resulting impulse, time and location. Impulse and particle data are
then written to multiple VTK files per frame. Samples of the smoothed color field (common
name in the literature for the field having the value 1 at particle locations and 0
everywhere else) on the regular cell lattice are also exported, which are then used to
reconstruct the fluid surface as an isosurface of that field. At the end of the simulation
a cumulative impulse heatmap along with the scaled and docked terrain model is also
provided.

\section{Results}

\paragraph{} Multiple simulations were carried out using different models of urban
coastline, in order to gain a significant and diverse dataset of impulses exerted over the
duration of the impact. Tsunamis are vastly different from the usual wind-induced sea
waves in that they have far longer wavelength and carry much greater total energy,
appearing as a rapidly rising tide instead of breaking waves. Accounting for these facts,
we chose to represent the tsunami wave as a water volume invading the coastline with an
initial velocity.

\begin{figure}[h!]
  \includegraphics[width=\textwidth]{figures/paraview.png}
  \caption{Example visualization of impact data in Paraview\textsuperscript{TM}. Fluid
    surface is reconstructed as an isosurface of the color field, while impulses on the
    selected region are vi\-su\-a\-li\-zed as points on the 3D terrain model and in a
    histogram grouped by their magnitude.}
  \label{fig:visualization}
\end{figure}

\begin{figure}[h!]
  \includegraphics[width=.33\textwidth]{figures/impulse-heatmap-0.png}
  \includegraphics[width=.32\textwidth]{figures/impulse-heatmap-1.png}
  \includegraphics[width=.33\textwidth]{figures/impulse-heatmap-2.png}
  \caption{Tsunami impulse heatmaps for three different urban models visualized under
    Paraview\textsuperscript{TM}. Most of the energy of the waves is absorbed by the first
    obstacle in their way.}
  \label{fig:impulse-heatmaps}
\end{figure}

\paragraph{} Figure \ref{fig:visualization} shows a typical simulation result with the aid
of Paraview\textsuperscript{TM}, where the reconstructed wave surface and impulses exerted
on a selected region are visualized along the 3D terrain model and plotted on a
histogram. Figure \ref{fig:impulse-heatmaps} shows impulse heatmaps generated by
simulation of tsunami impact upon three different urban terrain models. The base ground
width and depth of the models are in the order of 100 metres and the fluid is discretized
into 80k particles, having an initial velocity of 10m/s. Simulation time for each model
was about 40 minutes on a x86\_64 GNU/Linux desktop computer with an
Intel\textsuperscript{TM} i7-3770 processor and 8GB of main memory, accounting for 5
seconds of simulated time.

\paragraph{} A comparison of the resulting impulse heatmap from the same tsunami impact
upon an exposed coastline and one with a protective seawall is shown in Figure
\ref{fig:impulse-heatmaps-seawall}. These heatmaps confirm a well-documented observation
in the literature, i.e. that tsunami energy is absorbed mostly by the first obstacle in
its way. This has been noted by Danielsen et al. \cite{danielsen2005asian} and Kathiresan
and Narayanasamy \cite{kathiresan2005601}, who both emphasized the protective role coastal
vegetation (mangrove forests) played in mitigating the impact effects of the 2004 Indian
Ocean tsunami. The same conclusion has been reached by Yanagisawa et
al. \cite{yanagisawa200927} through theoretical approximation of the phenomenon backed by
relevant field data and measurements. Seawalls have been constructed in high-risk regions
as a common countermeasure against tsunami hazards. Although the waves may be so large as
to overtop such barriers, these cases are somewhat rare, making seawalls an effective
first line of defense.

\begin{figure}
  \includegraphics[width=\textwidth]{figures/if-city0-free.png}
  \includegraphics[width=\textwidth]{figures/if-city0-seawall.png}
  \caption{Comparison between the impulse heatmap of a tsunami hit on the same urban
    coastline model, with (up) and without (down) the protection from a seawall. Seawalls
    are commonly found in high-risk areas, as they reduce impact effects significantly.}
  \label{fig:impulse-heatmaps-seawall}
\end{figure}

\paragraph{} The performance of the simulation program was satisfactory, there is however
a substantial margin for improvement. Figure \ref{fig:performance} shows a plot of the
mean computation time per internal timestep for variable fluid resolution. It is important
to note that a higher number of simulation particles imposes a shorter internal timestep
due to their smaller effective radius. These measurements are also representative of the
worst-case scenario, as due to the initial fluid conditions the particles are closely
packed in a single body of fluid, thus having maximum number of mutual interactions and
reducing the performance gains from our spatial hashing data structure. About 75\% of
runtime is spent in single-threaded code consisting mostly of Bullet and I/O operations,
while the rest corresponds to multi-threaded code of our SPH engine.

\begin{figure}
  \includegraphics[width=\textwidth]{figures/performance.pdf}
  \caption{Plot showing the mean simulation time (in seconds) per internal timestep for
    the first 10 output frames (each frame was set to 0.05 seconds of simulated time), for
    10k to 80k particles (continuous line represents $O(n log_2n)$ growth).}
  \label{fig:performance}
\end{figure}

\paragraph{} Although our framework provides rich and detailed data, a quantitative
comparison to either field measurements or results of other methods utilizing different
versions of the SPH formulations couldn't be reliably made. We haven't yet implemented a
way to configure the wave's initial conditions and coastline terrain to model the impact
according to predictions of large scale simulations or data from past
incidents. Qualitatively, our method seems to compare well to similar codes utilizing SPH
to simulate small scale tsunami interactions with coastline structures. Other methods
usually make use of SPH enhancements originating from the CFD literature, which aim to
improve some aspect of the flow (for example incompressibility, regarding WCSPH). In this
particular application domain though, boundary interactions can be arguably considered
more important than the flow itself. Our approach, enriching SPH with direct positional
constraint solving, shows promise as an efficient way for dealing explicitly and exactly
with momentum preservation and detailed fluid-structure impulse recording.

\section{Conclusions}

\paragraph{} We presented the theoretical background and implementation of an enriched SPH
framework, whose main purpose was the faithful simulation and detailed recording of the
forces exerted upon the coastline during a tsunami impact. The chosen output file format
lends itself to rich visualizations, allowing for a quick overview of complex simulation
data. Multiple simulations have been carried out, resulting in interesting visualizations
of impact data and evaluation of defense mechanisms. Furthermore, detailed impulse logs
are provided, which can then be processed to extract relevant high level
information. Possible future extensions could be the incorporation of wider, dynamic
terrain models, including nearby seabed, the adjustment of the fluid initial conditions to
match those predicted by large scale shallow water or Boussinesq models and performance
optimizations in the implementation towards a more integrated dataflow scheme.

% ---- Bibliography ----
\begin{thebibliography}{14}

\bibitem{akinci2012versatile} Akinci, N., Ihmsen, M., Akinci, G., Solenthaler, B.,
  Teschner, M.: Versatile rigid-fluid coupling for incompressible SPH. ACM Transactions on
  Graphics (TOG) 31(4), 62:1--62:8 (2012)

\bibitem{becker2007weakly} Becker, M., Teschner, M.: Weakly compressible SPH for free
  surface flows. In Proceedings of the 2007 ACM SIGGRAPH/Eurographics symposium on
  Computer animation, 209--217 (2007)

\bibitem{dalrymple2006} Dalrymple, R. A., Rogers, B. D.: Numerical modeling of water waves
  with the SPH method. Coastal engineering 53(2), 141--147 (2006)

\bibitem{danielsen2005asian} Danielsen, F., S{\o}rensen, M. K., Olwig, M. F., Selvam, V.,
  Parish, F., Burgess, N. D., Hiraishi, T., Karunagaran, V. M., Rasmussen, M. S., Hansen,
  L. B. and Quarto, A.: The Asian tsunami: a protective role for coastal
  vegetation. Science(Washington) 310(5748), 643 (2005)

\bibitem{debroux2001three} Debroux, F., Prakash, M., Cleary, P.: Three-dimensional
  modelling of a tsunami interacting with real topographical coastline using Smoothed
  Particle Hydrodynamics. Proceedings of the 14th Australasian Fluid Mechanics Conference,
  Adelaide, Australia, 311--314 (2001)

\bibitem{desbrun1996smoothed} Desbrun M., Cani, M. P.: Smoothed particles: A new paradigm
  for animating highly deformable bodies. In Computer Animation and Simulation ’96
  (Proceedings of EG Workshop on Animation and Simulation), 61--76 (1996)

\bibitem{dominguez2011} Dom\'{i}nguez, J. M., Crespo, A. J., G\'{o}mez-‐Gesteira, M.,
  Marongiu, J. C.: Neighbour lists in smoothed particle hydrodynamics. International
  Journal for Numerical Methods in Fluids 67(12), 2026--2042 (2011)

\bibitem{gingold1977375} Gingold, R. A., Monaghan, J. J.: Smoothed particle hydrodynamics:
  theory and application to non-spherical stars. Monthly notices of the royal astronomical
  society 181(3), 375--389 (1977)

\bibitem{gomez2004} G\'{o}mez-Gesteira, M., Dalrymple, R. A.: Using a three-dimensional
  smoothed particle hydrodynamics method for wave impact on a tall structure. Journal of
  Waterway, Port, Coastal and Ocean Engineering 130(2), 63--69 (2004)

\bibitem{ihmsen2014implicit} Ihmsen, M., Cornelis, J., Solenthaler, B., Horvath, C.,
  Teschner, M.: Implicit incompressible SPH. Visualization and Computer Graphics, IEEE
  Transactions on 20(3), 426--435 (2014)

\bibitem{ioualalen2007} Ioualalen, M., Asavanant, J., Kaewbanjak, N., Grilli, S. T.,
  Kirby, J. T., Watts, P.: Modeling the 26 December 2004 Indian Ocean tsunami: Case study
  of impact in Thailand. Journal of Geophysical Research: Oceans 112(C7), (2007)

\bibitem{kakinuma2008} Kakinuma, T.: 3D numerical simulation of tsunami runup onto a
  complex beach. Advanced numerical models for simulating tsunami waves and runup,
  255--260 (2008)

\bibitem{kathiresan2005601} Kathiresan, K., Rajendran, N.: Coastal mangrove forests
  mitigated tsunami. Estuarine, Coastal and Shelf Science 65(3), 601--606 (2005)

\bibitem{lucy19771013} Lucy, L. B.: A numerical approach to the testing of the fission
  hypothesis. The astronomical journal 82, 1013--1024 (1977)

\bibitem{macklin2013position} Macklin, M., M\"{u}ller, M.: Position Based Fluids. ACM
  Trans. Graph. 32(4), 104:1--104:12 (2013)

\bibitem{muller2003particle} M\"{u}ller, M., Charypar, D., Gross, M.: Particle-based fluid
  simulation for interactive applications. Proceedings of the 2003 ACM
  SIGGRAPH/Eurographics symposium on Computer animation, 154--159 (2003)

\bibitem{samaras2015} Samaras, A. G., Karambas, T. V., Archetti, R.: Simulation of tsunami
  generation, propagation and coastal inundation in the Eastern Mediterranean. Ocean
  Science 11(4), 643--655 (2015)

\bibitem{solenthaler2009predictive} Solenthaler, B., Pajarola, R.: Predictive-corrective
  incompressible SPH. In ACM transactions on graphics (TOG) 28(3) 40:1--40:6 (2009)

\bibitem{st-germain2014} St-Germain, P., Nistor, I., Townsend, R., Shibayama, T.:
  Smoothed-particle hydrodynamics numerical modeling of structures impacted by tsunami
  bores. Journal of Waterway, Port, Coastal and Ocean Engineering 140(1), 66--81 (2014)

\bibitem{wang2007} Wang, X., Liu, P. L. F.: Numerical simulations of the 2004 Indian Ocean
  tsunamis—coastal effects. Journal of Earthquake and Tsunami 1(03), 273--297 (2007)

\bibitem{yanagisawa200927} Yanagisawa, H., Koshimura, S., Goto, K., Miyagi, T., Imamura,
  F., Ruangrassamee, A., Tanavud, C.: The reduction effects of mangrove forest on a
  tsunami based on field surveys at Pakarang Cape, Thailand and numerical
  analysis. Estuarine, Coastal and Shelf Science 81(1), 27--37 (2009)

\end{thebibliography}

\end{document}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: t
%%% End: